Loop-Aware Optimizations in PyPy's Tracing JIT

### 2012-08-16 talk/iwtc11/paper.tex
new benchmark results and computer info. formatting is screwed

\begin{figure}
\begin{center}
{\smaller
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
 & CPython & Psyco & PyPy  & PyPy & GCC \\
 &         &       & no LP &      & -O3 \\
\hline
conv3(1e5) & 77.89 & 9.52 & 1.77  & 0.68 &  0.59 \\
\hline
conv3(1e6) & 77.15 & 9.58 & 1.69  & 0.77 &  0.74 \\
\hline
conv3x3(1000) & 233.54 & 125.40 & 0.57 & 0.27 & 0.25 \\
\hline
conv3x3(3) & 234.45 & 126.28 & 0.60 & 0.31 & 0.28 \\
\hline
conv5(1e5) & 122.54 & 16.67 & 1.86  & 1.05 &  0.65\\
\hline
conv5(1e6) & 125.77 & 16.80 & 1.92 & 1.09 &  0.80 \\
\hline
dilate3x3(1000) & 232.51 & 125.85 & 3.89 & 3.69 & 0.25 \\
\hline
sobel(1000) & 181.49 & 95.05 & 0.71 & 0.42 & 0.20 \\
\hline
sqrt(Fix16) & 744.35 & 421.65 & 3.93  & 2.14  & 0.96 \\
\hline
sqrt(float) & 24.21 & 5.52 & 1.36 & 1.00 & 0.98\\
\hline
sqrt(int) & 20.84 & 1.78 & 2.26 & 1.82  & 0.80 \\
\hline
\hline
Variations & - & - & $\pm 0.03$ & $\pm 0.01$ & $\pm 0.01$ \\
\hline
\end{tabular}
}
\end{center}
\label{fig:benchmarks}
\caption{Benchmark Results in Seconds. Arrays of length $10^5$ and
  $10^6$ and matrixes of size $1000\times 1000$ and $1000000 \times
  3$ are used. The one used in each benchmark is indicated in
  the leftmost column. For the matrixes, only the number of rows are
  specified.} 
\end{figure}

@@

\revc{
Include the OS with your benchmark specifications.
}

Benchmarks were run on Intel i7 M620 @2.67GHz with 4M cache and 8G of RAM in
32bit mode.
The machine was otherwise unoccupied. We use the following software
for benchmarks:

>>>>

\begin{figure}
\begin{center}
{\smaller
\begin{tabular}{|l|r|r|r|r|r|r|r|}
\hline
 & CPython & PyPy  & PyPy & GCC & luajit & luajit \\
 &        & no LP &       & -O3 &        & no LP \\
FFT(1024, 32768) & 469.07 & 20.83 +- 0.02 & 12.73 +- 0.01 & - & 2.76 +- 0.01 & 4.42 +- 0.01\\
\hline
FFT(1024,32768) & - & - & - & 1.40 +- 0.04 & - & -\\
\hline
FFT(1048576, 2) & 58.93 & 4.12 +- 0.01 & 2.05 +- 0.00 & - & 1.08 +- 0.01 & 1.33 +- 0.01\\
\hline
FFT(1048576,2) & - & - & - & 0.83 +- 0.02 & - & -\\
\hline
LU(100, 4096) & 1974.14 & 32.22 +- 0.14 & 13.39 +- 0.03 & - & 1.52 +- 0.01 & 8.57 +- 0.01\\
\hline
LU(100,4096) & - & - & - & 1.33 +- 0.04 & - & -\\
\hline
LU(1000, 2) & 955.31 & 14.98 +- 0.22 & 5.99 +- 0.21 & - & 0.67 +- 0.01 & 3.99 +- 0.01\\
\hline
LU(1000,2) & - & - & - & 0.65 +- 0.04 & - & -\\
\hline
MonteCarlo(268435456) & 618.89 & 20.60 +- 0.05 & 15.33 +- 0.08 & 1.69 +- 0.05 & 2.82 +- 0.00 & 3.92 +- 0.01\\
\hline
SOR(100, 32768) & 1458.12 & 8.24 +- 0.00 & 2.66 +- 0.00 & - & 1.31 +- 0.01 & 2.02 +- 0.00\\
\hline
SOR(100,32768) & - & - & - & 1.76 +- 0.04 & - & -\\
\hline
SOR(1000, 256) & 1210.45 & 6.48 +- 0.00 & 2.10 +- 0.00 & - & 1.08 +- 0.01 & 1.63 +- 0.00\\
\hline
SOR(1000,256) & - & - & - & 1.49 +- 0.02 & - & -\\
\hline
SparseMatMult(1000, 5000, 262144) & 371.66 & 24.25 +- 0.04 & 16.52 +- 0.04 & - & - & -\\
\hline
SparseMatMult(1000,5000,262144) & - & - & - & 1.84 +- 0.03 & 4.53 +- 0.02 & 9.64 +- 0.02\\
\hline
SparseMatMult(100000, 1000000, 1024) & 236.93 & 17.01 +- 0.01 & 8.75 +- 0.08 & - & - & -\\
\hline
SparseMatMult(100000,1000000,1024) & - & - & - & 1.20 +- 0.03 & 2.42 +- 0.01 & 7.19 +- 0.01\\
\hline
conv3(100, nil) & - & - & - & - & 0.18 +- 0.00 & 0.70 +- 0.00\\
\hline
conv3(1000, nil) & - & - & - & - & 0.12 +- 0.01 & 0.67 +- 0.01\\
\hline
conv3(1e5) & - & - & - & 0.52 +- 0.04 & - & -\\
\hline
conv3(1e6) & - & - & - & 0.60 +- 0.03 & - & -\\
\hline
conv3(array(1e5)) & 50.14 & 1.09 +- 0.01 & 0.49 +- 0.01 & - & - & -\\
\hline
conv3(array(1e6)) & 49.20 & 1.13 +- 0.02 & 0.51 +- 0.00 & - & - & -\\
\hline
conv3x3(1000) & - & - & - & 0.17 +- 0.04 & - & -\\
\hline
conv3x3(1000, 1000) & - & - & - & - & 0.09 +- 0.01 & 0.49 +- 0.01\\
\hline
conv3x3(1000000, 3) & - & - & - & - & 0.13 +- 0.00 & 0.53 +- 0.00\\
\hline
conv3x3(3) & - & - & - & 0.19 +- 0.03 & - & -\\
\hline
conv3x3(Array2D(1000000x3)) & 139.81 & 0.70 +- 0.00 & 0.21 +- 0.00 & - & - & -\\
\hline
conv3x3(Array2D(1000x1000)) & 138.95 & 0.70 +- 0.00 & 0.20 +- 0.00 & - & - & -\\
\hline
conv5(100, nil) & - & - & - & - & 0.21 +- 0.01 & 0.87 +- 0.01\\
\hline
conv5(1000, nil) & - & - & - & - & 0.17 +- 0.01 & 0.84 +- 0.00\\
\hline
conv5(1e5) & - & - & - & 0.55 +- 0.02 & - & -\\
\hline
conv5(1e6) & - & - & - & 0.58 +- 0.03 & - & -\\
\hline
conv5(array(1e5)) & 74.65 & 1.22 +- 0.00 & 0.64 +- 0.00 & - & - & -\\
\hline
conv5(array(1e6)) & 77.94 & 1.26 +- 0.00 & 0.68 +- 0.01 & - & - & -\\
\hline
dilate3x3(1000) & - & - & - & 0.17 +- 0.03 & - & -\\
\hline
dilate3x3(1000, 1000) & - & - & - & - & 0.09 +- 0.00 & 0.48 +- 0.01\\
\hline
dilate3x3(Array2D(1000x1000)) & 137.52 & 4.35 +- 0.01 & 3.91 +- 0.02 & - & - & -\\
\hline
sobel(Array2D(1000x1000)) & 104.02 & 0.49 +- 0.00 & 0.21 +- 0.00 & 0.17 +- 0.03 & - & -\\
\hline
sobel\_magnitude(1000, 1000) & - & - & - & - & 0.24 +- 0.01 & 0.60 +- 0.01\\
\hline
sqrt(Fix16) & 463.46 & 5.12 +- 0.00 & 2.96 +- 0.00 & 1.34 +- 0.03 & 1.14 +- 0.00 & 12.80 +- 0.04\\
\hline
sqrt(float) & 14.99 & 1.37 +- 0.00 & 0.89 +- 0.00 & 0.85 +- 0.04 & 0.83 +- 0.01 & 1.06 +- 0.01\\
\hline
sqrt(int) & 13.91 & 3.22 +- 0.02 & 2.65 +- 0.00 & 1.25 +- 0.03 & 0.83 +- 0.01 & 1.06 +- 0.00\\
\hline

\end{tabular}
}
\end{center}
\label{fig:benchmarks}
\caption{Benchmark Results in Seconds. Arrays of length $10^5$ and
  $10^6$ and matrixes of size $1000\times 1000$ and $1000000 \times
  3$ are used. The one used in each benchmark is indicated in
  the leftmost column. For the matrixes, only the number of rows are
  specified.} 
\end{figure}

@@

Benchmarks were run on Intel Xeon X5680 @3.33GHz with 12M cache and 16G of RAM
using Ubuntu Linux 11.4 in 64bit mode.
The machine was otherwise unoccupied. We use the following software
for benchmarks:
